---
title: "New File"
author: "Vincent Morin"
date: "12/3/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = F,error=F,comment = F)
require(tidyverse)
require(tidytext)
require(lubridate)
require(ggthemes)
require(textdata)
require(topicmodels)
require(recipes)
require(caret)
```

```{r}
# Import Dataset
Data = read_csv("Data/2016_Green_Weather_Data.csv") 

```

This file will handle the data analysis portion of the Final Project. 

We'll begin by splitting the data into test and training datasets with equal proportions of 75% and 25%, respectively. 

Additionally, we'll ensure the training data proportion is correct by doing a quick calculation. 

```{r}
# Set seed for reproducibility. 
set.seed(333) 




# Indexing - used parse data into proportions. 
index = createDataPartition(Data$Avg_Tip, p=.75,list=F) 




# Use 75% of the data as training data 
train_data = Data[index,]




# Use 25% of the data as testing data. 
test_data = Data[-index,]




# Calculate proportion of training data
# relative to total data. 
# Should be ~ 0.75. 
round(nrow(train_data)/nrow(Data),3)

```

We'll begin by exploring our data. 

```{r}
# Use Glimpse to check dataset values. 
train_data %>% 
  summarize_all(class) %>% 
  glimpse()
```

At this point, our data is mostly numeric.  We should also check information on our date range. 

```{r}
# Print a brief summary of # of observations, 
# min date, and max date. 

train_data %>% summarize( 
  min_date = min(date), 
  max_date = max(date))

```

We'll also check to see if there are any missing values (NAs) in the dataset. 

```{r}
# Check for missing values among the numeric variables.
sum(is.na(train_data))
```

There are no missing values in the dataset. 

Next, we'll print an illustration (histograms) to see the distribution of our variables. 

```{r}
# Print histograms of distributions of variables. 
train_data_plot = train_data %>% 
  select_if(is.numeric) %>% 
  gather(var,val) %>% 
  ggplot(aes(val,group=var)) + 
  geom_histogram(bins = 30) +
  facet_wrap(~var,scales="free",ncol=2)




# Save illustration to a png file, "train_data_plot.png"
ggsave(
  train_data_plot, 
  filename = "Images/train_data_plot.png", 
  device = "png",
  width = 25, 
  height = 50, 
  limitsize = FALSE, 
  dpi = 300 )




# Print results
print(train_data_plot)
```

Most of the data is distributed relatively normally. 

A unique example of an exception to this is: Avg_Extra. 

```{r}
# Print histogram of Avg_Extra. 
Avg_Extra_plot = train_data %>% 
  select(Avg_Extra) %>% 
  gather(var,val) %>% 
  ggplot(aes(val,group=var)) + 
  geom_histogram(bins = 30) +
  facet_wrap(~var,scales="free",ncol=2)




# Save illustration to a png file, "Avg_Extra_plot"
ggsave(
  Avg_Extra_plot, 
  filename = "Images/Avg_Extra_plot.png", 
  device = "png",
  width = 25, 
  height = 50, 
  limitsize = FALSE, 
  dpi = 300 )




# Print results
print(Avg_Extra_plot)
```

Next, we'll consider additional predictor variables to derive from the data. 

To start, we'll utilize what information we have on the date, to create a few new variables:

- Day of the week,
- Day of the month,
- Week of year,
- Weekday - True or False, 
- Month.

This will be done using the lubridate package. 

```{r}
train_data1 <- train_data %>%
  mutate(Day_of_Week = 
           wday(train_data$date),
         Day_of_Month = 
           month(train_data$date),
         Week = 
           week(train_data$date), 
         Month = 
           month(train_data$date), 
         Weekday = ifelse(
           Day_of_Week > 1 & 
             Day_of_Week < 7, 
           "TRUE", "FALSE")
         )
         



# Use Glimpse to check dataset values. 
train_data1 %>% head()
```


```{r}
# Average Pickups Per Day
Average_Trips_Day_plot <- train_data1[, list(mean_trips_by_day = mean("Total_Trips")), by = Day_of_Week] %>%
  ggplot(aes(x = Day_of_Week, y = mean_trips_by_day)) +
  geom_bar(stat = "identity", fill = color) +
  labs(x = "Day of the Week", y = "Average Trips", title = "Average Trips by Day of Week")


print(Average_Trips_Day_plot)
```

