---
title: "Project Report"
author: "Vincent Morin"
date: "12/2/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = F,error=F,comment = F)
require(tidyverse)
require(tidytext)
require(lubridate)
require(ggthemes)
require(textdata)
require(topicmodels)
require(recipes)
require(caret)
```

```{r}
# Import Dataset
GreenTdata = read_csv("Data/2016_Green_Taxi_Trip_Data.csv") 
```

## Introduction

This analysis will attempt to comprehend the effects of weather on tips per taxi ride in New York City. 

With attendtion to the time-constraint associated with this project, the depth of our variable observations will be reduced, cleaned, or otherwise condensed. 

**Overview of Data**

We'll begin by taking a high-level overview to get a better idea of what our data looks like: our variables (columns) and a handful of observations (rows). 


```{r}
head(GreenTdata)

```


Because this is such a large dataset, we'll consider extranneous high-level material to remove.  Considerations to removed columns include: uncertainty/unknown descriptor, constant value, and location.  Sensitive to the constraint of the project, we will consider all trips to endure the weather recorded out of the Central Park weather station, and therefore remove columns related to location.   

**Trimming** 

We'll begin by removing several columns which have constant values, unknown descriptors, or are otherwise irrelevant to the analysis.  Some columns which were removed include pickup and dropoff locations

These include:
- VendorID
- Store_and_fwd_flag
- RateCodeID
- Pickup_longitude
- Pickup_latitude
- Dropoff_longitutde
- Dropoff_latitude
- Ehail_fee
- Trip_type
- Payment_type
- PULocationID
- DOLocationID


```{r}
# Remove several variables not relevant to goal. 
GreenTdata2 <- select(GreenTdata, -VendorID, -Store_and_fwd_flag, 
                      -RateCodeID, -Pickup_longitude, -Pickup_latitude, 
                      -Dropoff_longitude, -Dropoff_latitude, -Ehail_fee, 
                      -Trip_type, -Payment_type, -PULocationID, 
                      -DOLocationID)




head(GreenTdata2)
```

We'll also remove all rows without tips. 

```{r}
# filter out rows where observation in variable Tip_amount
# is equal to 0
GreenTips <- GreenTdata2[GreenTdata2$Tip_amount != 0, ]



```


**Clean Data**

Change column type for dates from character to date type. 

```{r}
GreenTips$lpep_pickup_datetime <- mdy_hms(GreenTips$lpep_pickup_datetime)



GreenTips$Lpep_dropoff_datetime <- mdy_hms(
  GreenTips$Lpep_dropoff_datetime)



head(GreenTips)
```

Include only half of the year. 


```{r}
GreenDate <- GreenTips %>% mutate(
  date = as.Date(GreenTips$lpep_pickup_datetime))




GreenData <- GreenDate %>% filter(
  GreenDate$date < "2016-7-1")




#GreenJanJun <- subset(GreenTips,  lpep_pickup_datetime > as.Date("2016-07-01"))

head(GreenData)
tail(GreenData)
```
Drop date variable

```{r}
GreenData <- GreenData %>% select(-date)
```


```{r}
# Save new datafile

GreenData %>% write_csv(path = "Data/2016_Green_Taxi_Data.csv")



```




